---
title: "Assignment 2:Electricity Market Prices"
author: "Enes Kacar"
date: "JULY 2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(tidyr)
library(tinytex)
library(lubridate)
library(ggplot2)
library(dplyr)
```
# Dataset

###  Terminology

PTF = Market Clearing Price

SMF = System Marginal Price

PDF = Positive Unbalance Price

NDF = Negative Unbalance Price

For more detailed information, you can visit the [EPİAŞ](https://www.epias.com.tr/spot-elektrik-piyasasi) website.

### Reading Data
```{r}
ptfsmf <- read_excel('ptf-smf.xls')
```
```{r}
glimpse(ptfsmf)
```
### Editting Data
```{r}
# Arrangement of the names of the positive and negative imbalance columns
colnames(ptfsmf)[4] = "PDF"
colnames(ptfsmf)[5] = "NDF"

# Converting the date and time to POSIX format
ptfsmf <- ptfsmf %>% mutate(Tarih_Yeni = 
                              as.POSIXct(ptfsmf$Tarih,format="%d.%m.%y %H:%M", "GMT")) %>%
                     select(Tarih_Yeni, PTF, SMF, PDF, NDF) %>%
                     rename(Tarih = Tarih_Yeni)

# Converting the remaining columns to numeric class
ptfsmf[,c(2:5)] <- lapply(ptfsmf[,c(2:5)],
                          function(x) as.numeric(gsub(",", ".", gsub("\\.", "", x))))
```
# 1.Weekly Analysis

### Data preparition for weekly analysis
```{r}
weekly_analysis <- ptfsmf %>%
  select(Tarih,PTF,SMF,PDF,NDF) %>%
  mutate(Week=as.factor(week(Tarih)))%>%
  pivot_longer(cols = c(PTF,SMF,PDF,NDF),names_to = 'names',values_to='values')%>%
  group_by(Week,names)%>%
  summarise(sum_price=sum(values),mean_price=mean(values),max_price = max(values),min_price=min(values),.groups = 'drop')%>%
  transform(Week=as.numeric(Week))
glimpse(weekly_analysis)
```

### Visualiation

```{r}
ggplot(weekly_analysis,aes(x=Week, y = mean_price, fill = names)) +
  geom_col(position="dodge2")+
  labs(y=" Weekly Average of Prices")+
  theme_minimal()
  
```
```{r}
ggplot(weekly_analysis,aes(x=Week, y = max_price, fill = names)) +
  geom_col(position='dodge2')+
  labs(y=" Weekly Max Prices")+
  theme_minimal()
```


```{r}
ggplot(weekly_analysis,aes(x=Week, y = min_price, fill = names)) +
  geom_col(position="dodge2")+
  labs(y=" Weekly Minimum Prices")+
  theme_minimal()
```



# 2. Daily Analysis

### Data preparition for hourly analysis

```{r}
daily_analysis <- ptfsmf %>%
  select(Tarih,PTF,SMF,PDF,NDF) %>%
  mutate(Day=as.factor(day(Tarih)))%>%
  pivot_longer(cols = c(PTF,SMF,PDF,NDF),names_to = 'names',values_to='values')%>%
  group_by(Day,names)%>%
  summarise(sum_price=sum(values),mean_price=mean(values),max_price = max(values),min_price=min(values),.groups = 'drop')%>%
  transform(Day=as.numeric(Day))
glimpse(daily_analysis)
```

```{r}
ggplot(daily_analysis, aes( Day,mean_price)) + 
  geom_point()+
  geom_smooth() +
    labs(
        x = "Days of July", # making the axis label blank will remove it
        y = "Price(mean)")
```


# 3.Hourly Analysis

### Data preparition for hourly analysis
```{r}
hourly_analysis <- ptfsmf %>%
  select(Tarih,PTF,SMF,PDF,NDF) %>%
  mutate(Hour=as.factor(hour(Tarih)))%>%
  pivot_longer(cols = c(PTF,SMF,PDF,NDF),names_to = 'names',values_to='values')%>%
  group_by(Hour,names)%>%
  summarise(sum_price=sum(values),mean_price=mean(values),max_price = max(values),min_price=min(values),.groups = 'drop')%>%
  transform(Hour=as.numeric(Hour))
glimpse(hourly_analysis)
```
### Visualization
```{r}
ggplot(hourly_analysis, aes(x = Hour, y = mean_price)) +
    geom_point() + 
    geom_smooth() +
    labs(title = "Hourly Analysis",
        x = "Days of June", 
        y = "Price(mean)")
```



#  PTF vs SMF

```{r}
ggplot(ptfsmf,aes(x=PTF,y=SMF)) + 
  geom_point()+ 
  geom_smooth() +
  labs(
        x = "PTF",
        y = "SMF")
```













